---
title: "Obj3_CN"
output: html_document
date: "2024-12-23"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## C & N Analyses

Set-up

```{r}
# load packages
library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
library(lme4)
library(lmerTest)
library(emmeans)
library(viridis)
library(MuMIn)
library(vegan)
library(stringr)

path <- "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/"
```

## Data Loading and Tidying ONLY DO ONCE

#### Merge all of the C and N tin plate runs into one dataframe

```{r}
    ## Do only once to tidy raw data for subsequent analyses

    #load data 
    CN2 <-read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/RAW/Whitman Muscettola CN 032824.csv")
    CN1 <-read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/RAW/Whitman Muscettola CN 051424.csv")
    CN3 <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/RAW/CN results for LAV022 Elm.csv")

    #add column with analysis date and run number
    CN2$Date <- 20240328
    CN1$Date <- 20240514
    CN3$Date <- 20250809
    CN2$Run <- 1
    CN1$Run <- 1
    CN3$Run <- 2

    #add Notes column to CN3
    CN3$Notes <- ""

    #merge two datasets from EA analysis runs 1 and 2 
    temp <- rbind(CN1, CN2)
    CN <- bind_rows(temp, CN3)
    rm(temp)

    #rename column X.N and X.C to percN and percC
    CN <- CN %>%
      rename(percN = X.N) %>%
      rename(percC = X.C)

    #add columns removing data for samples with percent below or above detection limit 
    CN <- CN %>%
    mutate(cleanN = ifelse(Notes=="N below detection limit", NA, ifelse(Notes=="CN below detection limit",NA,percN))) %>% 
      mutate(cleanC = ifelse(Notes=="C above detection limit",
    NA, ifelse(Notes=="CN below detection limit", NA, percC)))

    #separate out samples that were rerun 
    rerunsamples <- CN |> group_by(Sample.Name) |> filter(n()>1) |> ungroup()
    
    write.csv(rerunsamples, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CN_rerunsamples.csv")
    

    #what were the issues with the samples that needed to be rerun?
    temp <- CN %>% 
      group_by(Notes) %>% 
      summarise(count=n()) 
    rm(temp)

    write.csv(CN,"C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CN_allruns_merged.csv")
    

# the merged CN dataframe includes multiple runs of the same samples. This is either a) because the sample was accidentally ran twice, b) because the sample was run and at least one of the C or N values was outside of detection limits, c) the sample was run and did not return any data for C or N. This function finds the duplicates based on the Sample.Name. Then among the duplicates it will keep only one row - 1) ideally a row where both cleanC & clean N are numbers (i.e. not NA) or 2) keep a row where at least one of the two is a number
    
#NOTE: 2 errors in sample name was identified in the QA/QC - corrections were made to the csv and file was named "CN_allruns_merged_corrected.csv" this merged csv should be used for subsequent analyses.
    
    CN <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CN_allruns_merged_corrected.csv")


    CN_tidy <- CN %>%
  group_by(Sample.Name) %>%
  #create a "score" based on our preferences outlined above for cleanC & cleanN values
  mutate(score = case_when(
    !is.na(cleanC) & ! is.na(cleanN) ~2, #both present
    !is.na(cleanC) | !is.na(cleanN) ~1, #one present
    TRUE ~ 0 #both NA
  )) %>%
  #keep the rows with the highest score
  filter(score == max(score)) %>%
  # in case of ties, pick the first occurrence
  slice(1) %>%
  ungroup() %>%
  select(!score) %>%
  #rename "Date" column to "CN Analysis Date" & "Weight" to "CN_Weight"
  rename(CN_Analysis_Date = Date, CN_Weight = Weight) %>%
  #select relevant columns to the metadata and subsequent analyses
  select(Sample.Name, CN_Analysis_Date, CN_Weight:cleanC)
    
    #replace " " in Sample.Name with "_" so that it matches the Sample Name in the metadata
    CN_tidy$Sample.Name <- gsub(" ", "_", CN_tidy$Sample.Name)
    
    write.csv(CN_tidy, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CN_merged_final.csv")
    
rm(CN1, CN2, CN3, temp, CN, rerunsamples)
```

#### Add C/N concentrations to metadata dataframe

```{r}
    #add C/N concentrations to metadata 
    metadata <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/03112024_LAVO22_Metadata.csv")
    
    CN_tidy <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CN_merged_final.csv")
    
      
    newmetadata <- left_join(metadata, CN_tidy[, 2:6], by=c("Full_Sample_Name" = "Sample.Name")) %>%
      mutate(., Plot_Horizon = paste(Plot_Name, Horizon, sep="_"))
      #includes a column for both the plot and horizon (this is useful when looking at similarity of soil core characteristics within the same plot)
    
    #save metadata 
    write.csv(newmetadata,"C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/09202025_LAVO22_Metadata.csv")
    rm(temp, CN1, CN2, CN, metadata)
```

### Tidy up metadata

```{r}
    
    newmetadata <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/09202025_LAVO22_Metadata.csv")

    #filter data so no duplicates and remove columns pertinent for sequencing 
    df <- newmetadata %>%
      filter(startsWith(Run_Submission_Name, "1")) %>%
      #removes 2nd sequencing run 
      filter(!endsWith(Run_Submission_Name, "2")) %>% 
      #removes duplicate samples from sequencing 
      select(-1, -(3:7), -(15:17), -(23:27))
      #removes extraneous columns that are related to DNA sequencing

    #set column types 
    df$Unit <- factor(df$Unit)
    df$SampleSev <- factor(df$SampleSev, levels = c(1, 2, 3, 4))
    
    
    #filter data to remove unburned unit 
    df.noUB <- filter(df, Unit!="UB")

    #save dataframes
    write.csv(df.noUB, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/LAVO22_CN_noUB_metadata.csv")
    
    write.csv(df, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/LAVO22_CN_allunits_metadata.csv")
```    

### Which samples are missing Carbon and/or Nitrogen data?
```{r}

    #which samples are missing either C or N or both?
    CN_allruns_merged <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CN_allruns_merged_corrected.csv")
    
    missingdata <- CN_allruns_merged %>%
  group_by(Sample.Name) %>%
  #create a "score" based on our preferences outlined above for cleanC & cleanN values
  mutate(score = case_when(
    !is.na(cleanC) & ! is.na(cleanN) ~2, #both present
    !is.na(cleanC) | !is.na(cleanN) ~1, #one present
    TRUE ~ 0 #both NA
  )) %>%
  #keep the rows with the min score (in cases where there are more than 1 run per sample)
  mutate(maxscore = max(score)) %>%
  filter(!maxscore == 2) %>%
  select(-score, -maxscore, -X)

# save missing data
    write.csv(missingdata, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/20250920_missingCN_samples.csv")
    #when compiling the list of samples to rerun, I had only included those from the 5/14/2025 run, I missed those that were lacking data from the 3/28/2025 run. This resulted in a total of 20 samples that have at least 1 C/N value missing - 19 missing N data and 1 missing both N and C data. 18 missing values come from the Butte Lake unit, 1 missing from the Hole unit and 1 missing from the unburned unit. 
    
    
    #which samples were never run for C/N analyses?
    missing <- df %>%
      filter(is.na(cleanC) & is.na(cleanN) & is.na(CN_Analysis_Date)) %>%
      select(Full_Sample_Name, CN_Analysis_Date:cleanC) %>%
      mutate(Notes=NA, Run.ID=NA, percN=NA, percC=NA, Run=NA) %>%
      rename(Sample.Name=Full_Sample_Name, Date=CN_Analysis_Date, Weight=CN_Weight)
    
    #merge with the dataframe previously created for missing data
    missingdata <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/20250920_missingCN_samples.csv")
    missingdata <- select(missingdata, -X)
    
    all_missing <- rbind(missing, missingdata)
    
    #save all missing data as its own dataframe
    write.csv(all_missing, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/allmissingCN_samples.csv")
```

```         
</div>
```

```{r}
#load tidied C/N data 
CN.all <- read.csv("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/LAVO22_CN_allunits_metadata.csv")
#remove extraneous first column
CN.all <- select(CN.all, -1)
df.noUB <- CN.all %>%
  filter(Unit!="UB")

#make sure columns are correct data type
df.noUB$Unit <- factor(df.noUB$Unit)
df.noUB$SampleSev <- factor(df.noUB$SampleSev, levels = c(1, 2, 3, 4))
df.noUB$Horizon <- factor(df.noUB$Horizon)
```

## Conversion of concentration to stock

```{r}
df <- read.csv(paste0(path, "Data/R/CarbonNitrogen/CN_merged_final.csv"))

df <- df %>%
  mutate(Plot_Name=str_sub(Sample.Name, 1, -4)) %>%
  #there appears to be a duplicate sample (BL PIJE 04 AM) let's remove one of the reads for a cleaner dataframe. When made into final code and dataframe this will need to be addressed during the "Tidy up metadata" section of code. 
  filter(!endsWith(Plot_Name,"0")) %>%
  mutate(Plot_Core=paste0(Plot_Name, str_sub(Sample.Name, -2, -2)))

### summarize CN for the average of each plot
CN.df.summary <

df.noUB$CStock <- df.noUb$cleanC * df.noUb$bd * (1- df.noUb$CF) *5
df.noUB$NStock <- df.noUb$cleanC * df.noUb$bd * (1-df.noUb$CF) *5
df.noUB$CNratio <- df.noUb$cleanC/df.noUb$cleanN
```

## Data Visualization

### % C & % N

```{r}
#some data exploration - let's look at % C/N for all units
N.allunits <- ggplot(df, aes(Unit, cleanN)) + geom_point() + theme_bw() + facet_wrap(~Horizon, scales = "free_y", labeller = as_labeller(c("M"="Mineral", "O"="Organic"))) + geom_boxplot() + theme_bw() + scale_y_continuous(name="% Nitrogen") + scale_x_discrete(name="Unit", labels=c("Butte Lake", "Hole", "Unburned", "Warner Valley"))
N.allunits

ggsave("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Figures/CN/percN_allunits.png", plot = N.allunits, width = 10, height = 8, units = "in", dpi = 300)

C.allunits <- ggplot(df, aes(Unit, cleanC)) + geom_point() + theme_bw() + facet_wrap(~Horizon, scales = "free_y", labeller = as_labeller(c("M"="Mineral", "O"="Organic"))) + geom_boxplot() + theme_bw() + scale_y_continuous(name="% Carbon") + scale_x_discrete(name="Unit", labels=c("Butte Lake", "Hole", "Unburned", "Warner Valley"))
C.allunits

ggsave("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Figures/CN/percC_allunits.png", plot = C.allunits, width = 10, height = 8, units = "in", dpi = 300)

```

```{r}
#visualize Carbon data by treatment faceted by unit and horizon
C.firehistory <- ggplot(df.noUB, aes(Trt, cleanC, fill=Trt)) 
C.firehistory <- C.firehistory + geom_boxplot() + theme_bw() + facet_grid(Horizon~Unit, scales = "free_y", labeller = as_labeller(c("M"="Mineral", "O"="Organic", "BL"="Butte Lake", "H"="Hole", "WV"="Warner Valley"))) 
C.firehistory <- C.firehistory + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
C.firehistory <- C.firehistory + scale_y_continuous(name="% Carbon") 
C.firehistory <- C.firehistory + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire"))
C.firehistory

#save plot as a png file
ggsave("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Figures/CN/percC_bytreatmentandunit.png", plot = C.firehistory, width = 10, height = 8, units = "in", dpi = 300)

#visualize Nitrogen data by treatment faceted by unit and horizon
N.firehistory <- ggplot(df.noUB, aes(Trt, cleanN, fill=Trt)) 
N.firehistory <- N.firehistory + geom_boxplot() + theme_bw() + facet_grid(Horizon~Unit, scales="free_y", labeller=as_labeller(c("M"="Mineral", "O"="Organic", "BL"="Butte Lake", "H"="Hole", "WV"="Warner Valley"))) 
N.firehistory <- N.firehistory + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
N.firehistory <- N.firehistory + scale_y_continuous(name="% Nitrogen") 
N.firehistory <- N.firehistory + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire"))
N.firehistory

#save plot as a png file
ggsave("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Figures/CN/percN_bytreatmentandunit.png", plot = N.firehistory, width = 10, height = 8, units = "in", dpi = 300)
```

```{r}
# % Carbon by burn severity faceted by unit and horizon 
C.burnsev <- ggplot(df.noUB, aes(SampleSev, cleanC, fill=SampleSev)) 
C.burnsev <- C.burnsev + geom_boxplot() + theme_bw() + facet_grid(Horizon~unit, scales = "free_y", labeller = as_labeller(c("M" = "Mineral", "O" = "Organic", "Butte Lake" = "Butte Lake", "Hole" = "Hole", "Warner Valley" = "Warner Valley"))) 
C.burnsev <- C.burnsev + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
C.burnsev <- C.burnsev + scale_y_continuous(name="% Carbon") 
C.burnsev <- C.burnsev + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
C.burnsev

#save plot as a png file
ggsave("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Figures/CN/percC_byburnseverityandunit.png", plot = C.burnsev, width = 10, height = 8, units = "in", dpi = 300)

# % Nitrogen by burn severity faceted by unit and horizon
N.burnsev <- ggplot(df.noUB, aes(SampleSev, cleanN, fill=SampleSev)) 
N.burnsev <- N.burnsev + geom_boxplot() + theme_bw() + facet_grid(Horizon~unit, scales = "free_y", labeller = as_labeller(c("M" = "Mineral", "O" = "Organic", "Butte Lake" = "Butte Lake", "Hole" = "Hole", "Warner Valley" = "Warner Valley"))) 
N.burnsev <- N.burnsev + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
N.burnsev <- N.burnsev + scale_y_continuous(name="% Nitrogen") 
N.burnsev <- N.burnsev + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
N.burnsev

#save plot as a png file
ggsave("C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Figures/CN/percN_byburnseverityandunit.png", plot = N.burnsev, width = 10, height = 8, units = "in", dpi = 300)

```

### C & N Stocks; Stopped here 9/29/2025

```{r}
#visualize C/N data by treatment faceted by unit and horizon
Cstock.firehistory.M <- ggplot(df.noUB, aes(Trt, Cstock, fill=Trt)) 
Cstock.firehistory.M <- Cstock.firehistory.M + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Cstock.firehistory.M <- Cstock.firehistory.M + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
Cstock.firehistory.M <- Cstock.firehistory.M + scale_y_continuous(name="Carbon Stock (mg C/cm^2 soil)") 
Cstock.firehistory.M <- Cstock.firehistory.M + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire")) 
Cstock.firehistory.M

Cstock.firehistory.O <- ggplot(df.noUB[df.noUB$Horizon=="O", ], aes(Trt, CStock, fill=Trt)) 
Cstock.firehistory.O <- Cstock.firehistory.O + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Cstock.firehistory.O <- Cstock.firehistory.O + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
Cstock.firehistory.O <- Cstock.firehistory.O + scale_y_continuous(name="Carbon Stock (mg C/cm^2 soil)") 
Cstock.firehistory.O <- Cstock.firehistory.O + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire")) 
Cstock.firehistory.O

Nstock.firehistory.M <- ggplot(df.noUB[df.noUB$Horizon=="M", ], aes(Trt, NStock, fill=Trt)) 
Nstock.firehistory.M <- Nstock.firehistory.M + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Nstock.firehistory.M <- Nstock.firehistory.M + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
Nstock.firehistory.M <- Nstock.firehistory.M + scale_y_continuous(name="Nitrogen Stock (mg N/cm^2 soil)") 
Nstock.firehistory.M <- Nstock.firehistory.M + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire")) 
Nstock.firehistory.M

Nstock.firehistory.O <- ggplot(df.noUB[df.noUB$Horizon=="O", ], aes(Trt, NStock, fill=Trt)) 
Nstock.firehistory.O <- Nstock.firehistory.O + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Nstock.firehistory.O <- Nstock.firehistory.O + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
Nstock.firehistory.O <- Nstock.firehistory.O + scale_y_continuous(name="Nitrogen Stock (mg N/cm^2 soil)") 
Nstock.firehistory.O <- Nstock.firehistory.O + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire")) 
Nstock.firehistory.O

ggplot(df.noUB, aes(Trt, CStock)) +geom_boxplot() + facet_grid(Horizon~unit)

ggplot(df.noUB, aes(Trt, NStock)) +geom_boxplot() + facet_grid(Horizon~unit)
```

```{r}
#C/N data by burn severity by unit and horizon 
Cstock.burnsev.M <- ggplot(df.noUB[df.noUB$Horizon=="M", ], aes(SampleSev, CStock, fill=SampleSev)) 
Cstock.burnsev.M <- Cstock.burnsev.M + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Cstock.burnsev.M <- Cstock.burnsev.M + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
Cstock.burnsev.M <- Cstock.burnsev.M + scale_y_continuous(name="Carbon Stock (mg C/cm^2 soil)") 
Cstock.burnsev.M <- Cstock.burnsev.M + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
Cstock.burnsev.M

Cstock.burnsev.O <- ggplot(df.noUB[df.noUB$Horizon=="O", ], aes(SampleSev, CStock, fill=SampleSev)) 
Cstock.burnsev.O <- Cstock.burnsev.O + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Cstock.burnsev.O <- Cstock.burnsev.O + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
Cstock.burnsev.O <- Cstock.burnsev.O + scale_y_continuous(name="Carbon Stock (mg C/cm^2 soil)") 
Cstock.burnsev.O <- Cstock.burnsev.O + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
Cstock.burnsev.O

Nstock.burnsev.M <- ggplot(df.noUB[df.noUB$Horizon=="M", ], aes(SampleSev, NStock, fill=SampleSev)) 
Nstock.burnsev.M <- Nstock.burnsev.M + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Nstock.burnsev.M <- Nstock.burnsev.M + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
Nstock.burnsev.M <- Nstock.burnsev.M + scale_y_continuous(name="Nitrogen Stock (mg N/cm^2 soil)") 
Nstock.burnsev.M <- Nstock.burnsev.M + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
Nstock.burnsev.M

Nstock.burnsev.O <- ggplot(df.noUB[df.noUB$Horizon=="O", ], aes(SampleSev, NStock, fill=SampleSev)) 
Nstock.burnsev.O <- Nstock.burnsev.O + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
Nstock.burnsev.O <- Nstock.burnsev.O + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
Nstock.burnsev.O <- Nstock.burnsev.O + scale_y_continuous(name="Nitrogen Stock (mg N/cm^2 soil)") 
Nstock.burnsev.O <- Nstock.burnsev.O + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
Nstock.burnsev.O

ggplot(df.noUB, aes(SampleSev, percN)) + geom_point() + theme_bw() + facet_grid(Horizon~unit)
```

### C:N ratio

```{r}
#visualize C/N data by treatment faceted by unit and horizon
CN.firehistory.M <- ggplot(df.noUB[CN.noUB$Horizon=="M", ], aes(Trt, CNratio, fill=Trt)) 
CN.firehistory.M <- CN.firehistory.M + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
CN.firehistory.M <- CN.firehistory.M + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
CN.firehistory.M <- CN.firehistory.M + scale_y_continuous(name="C:N ratio") 
CN.firehistory.M <- CN.firehistory.M + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire")) 
CN.firehistory.M

CN.firehistory.O <- ggplot(df.noUB[df.noUB$Horizon=="O", ], aes(Trt, CNratio, fill=Trt)) 
CN.firehistory.O <- CN.firehistory.O + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
CN.firehistory.O <- CN.firehistory.O + scale_fill_manual(name="Fire History", values=c('#0097d3','#ffbb22'), labels = c("Less Fire", "More Fire"))
CN.firehistory.O <- CN.firehistory.O + scale_y_continuous(name="C:N ratio") 
CN.firehistory.O <- CN.firehistory.O + scale_x_discrete(name="Fire History", labels=c("Less Fire", "More Fire")) 
CN.firehistory.O

ggplot(df.noUB, aes(Trt, CNratio)) +geom_boxplot() + facet_grid(Horizon~unit)

ggplot(df.noUB, aes(Trt, CNratio)) +geom_boxplot() + facet_grid(Horizon~unit)
```

```{r}
#C/N data by burn severity by unit and horizon 
CN.burnsev.M <- ggplot(df.noUB[df.noUB$Horizon=="M", ], aes(SampleSev, CNratio, fill=SampleSev)) 
CN.burnsev.M <- CN.burnsev.M + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
CN.burnsev.M <- CN.burnsev.M + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
CN.burnsev.M <- CN.burnsev.M + scale_y_continuous(name="C:N ratio") 
CN.burnsev.M <- CN.burnsev.M + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
CN.burnsev.M

CN.burnsev.O <- ggplot(df.noUB[df.noUB$Horizon=="O", ], aes(SampleSev, CNratio, fill=SampleSev)) 
CN.burnsev.O <- CN.burnsev.O + geom_boxplot() + theme_bw() + facet_wrap(~unit) 
CN.burnsev.O <- CN.burnsev.O + scale_fill_manual(name="Burn Severity", values=c('#f5db4c', '#f98e09', '#bc3754', '#57106e'),labels = c("Unburned", "Low", "Moderate", "Severe")) 
CN.burnsev.O <- CN.burnsev.O + scale_y_continuous(name="C:N ratio") 
CN.burnsev.O <- CN.burnsev.O + scale_x_discrete(name="Burn Severity", labels=c("Unburned", "Low", "Moderate", "Severe"))
CN.burnsev.O


ggplot(df.noUB, aes(SampleSev, CNratio)) + geom_point() + theme_bw() + facet_grid(Horizon~unit)
```

## Statistics

linear mixed model with random effects of plot and core

### % C Model selection

```{r}
#Forwards model selection 
#Horizon 
mod.hor <- lmer(percC ~ Horizon + (1|Plot_Name/Rep), data=df.noUB) 
summary(mod.hor) 
anova(mod.hor)
#horizon is significant; p=2.2e-16

#Unit 
mod.unit <- lmer(percC~Unit + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) 
# \*\*\* p=1.6e-8

#Soil subgroup 
mod.soil <- lmer(percC~Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.soil) 
# \*\*\* p=0.0002097

#Sample severity 
mod.sev <-lmer(percC~SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
summary(mod.sev) 
anova(mod.sev) 
#\* p=0.030

#VegType 
mod.veg <- lmer(percC~Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.veg) 
#\*\*\* p=8.1e-6

#Trt 
mod.trt <- lmer(percC~Trt + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) 
# Trt not sig p=0.5

#unit and soil subgroup 
mod.u.sg <- lmer(percC~Unit+Soil_Subgroup + (1\|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.sg) 
#when accounting for unit, soil type is no longer a sig predictor

#unit and veg type 
mod.u.v <- lmer(percC~Unit+Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.v) 
#when accounting for unit, veg type is no longer a sig predictor

#unit and horizon 
mod.u.h <-lmer(percC~Unit+Horizon+(1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h) 
#both unit and horizon sig predictors

#unit, horizon, trt 
mod.u.h.t <- lmer(percC~Unit+Horizon+Trt +(1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t) 
#trt not sig predictor

#unit, horizon, samplesev 
mod.u.h.ss <- lmer(percC~Unit+Horizon+SampleSev+ (1|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) 
#burn severity is sig predictor

#unit, horizon, samplesev, trt 
mod.u.h.ss.t <- lmer(percC~Unit+Horizon+SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t) 
#trt still not sig predictor

##interactions 2-way 
mod.u.h.2.ss.t <- lmer(percC~Unit*Horizon+SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.t) 
#2-way interaction between unit and horizon (p=0.00048) 

mod.u.ss.2.h.t <- lmer(percC~Unit*SampleSev+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.ss.2.h.t) 
#no 2-way interaction of unit and samplesev 

mod.u.t.2.ss.h <- lmer(percC~Unit*Trt+SampleSev+Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.t.2.ss.h) 
#no Unit*Trt

mod.u.h.ss.2.t <- lmer(percC~Unit+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.2.t)
#Horizon*SampleSev 

mod.u.h.t.2.ss <- lmer(percC~Unit+Horizon*Trt+SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t.2.ss) #no Horizon*Trt 

mod.u.h.ss.t.2 <- lmer(percC~Unit+Horizon+SampleSev*Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t.2) #no SampleSev*Trt

mod.u.h.2.ss.2.t <- lmer(percC~Unit*Horizon+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.2.t) 
#unit*horizon and horizon*SampleSev sig

##interactions 3-way 
mod.u.h.ss.3.t <-lmer(percC~Unit*Horizon*SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.3.t) #no unit*horizon*samplesev

#from forward model selection appears that Unit*Horizon+Horizon*SampleSev+Trt is the best fit model
summary(mod.u.h.2.ss.2.t)

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) 
#model with lowest AIC: Unit*Horizon+Horizon*SampleSev+Trt (1831.3) 
#model with 2nd lowest AIC: Unit*Horizon*SampleSev+Trt (1834.3) 
#3rd lowest AIC: Unit\*Horizon+SampleSev+Trt (1834.7)

#%C model selected is Unit*Horizon+Horizon*SampleSev+Trt 
finalmod.C <- mod.u.h.2.ss.2.t 

rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)
```

### %N Model Selection

```{r}
#Forwards model selection #Horizon 
mod.hor <- lmer(percN ~ Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.hor) #horizon is significant; p=2.2e-16

#Unit 
mod.unit <- lmer(percN~Unit + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) # \*\*\* p=2e-13

#Soil subgroup 
mod.soil <- lmer(percN~Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.soil) # \*\*\* p=0.1.38e-6

#Sample severity 
mod.sev <-lmer(percN~SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.sev) #* p=0.04812

#VegType 
mod.veg <- lmer(percN~Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.veg) #\*\*\* p=2.0e-8

#Trt 
mod.trt <- lmer(percN~Trt + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) # Trt not sig p=0.608

#unit and soil subgroup 
mod.u.sg <- lmer(percN~Unit+Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.sg) #when accounting for unit, soil type is no longer a sig predictor

#unit and veg type 
mod.u.v <- lmer(percN~Unit+Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.v) #when accounting for unit, veg type is no longer a sig predictor

#unit and horizon 
mod.u.h <-lmer(percN~Unit+Horizon+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h) #both unit and horizon sig predictors

#unit, horizon, trt 
mod.u.h.t <- lmer(percN~Unit+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t) #trt not sig predictor

#unit, horizon, samplesev 
mod.u.h.ss <- lmer(percN~Unit+Horizon+SampleSev+ (1|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) #burn severity is sig predictor

#unit, horizon, samplesev, trt 
mod.u.h.ss.t <- lmer(percN~Unit+Horizon+SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t) #trt still not sig predictor

##interactions 2-way 
mod.u.h.2.ss.t <- lmer(percN~Unit*Horizon+SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.t) #2-way interaction between unit and horizon (p=1.88e-6) 

mod.u.ss.2.h.t <- lmer(percN~Unit*SampleSev+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.ss.2.h.t) #no 2-way interaction of unit and samplesev 

mod.u.t.2.ss.h <- lmer(percN~Unit*Trt+SampleSev+Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.t.2.ss.h) #no Unit*Trt

mod.u.h.ss.2.t <- lmer(percN~Unit+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.2.t) #Horizon*SampleSev 

mod.u.h.t.2.ss <- lmer(percN~Unit+Horizon*Trt+SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t.2.ss) #no Horizon*Trt 

mod.u.h.ss.t.2 <- lmer(percN~Unit+Horizon+SampleSev*Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t.2) #no SampleSev*Trt

mod.u.h.2.ss.2.t <- lmer(percN~Unit*Horizon+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.2.t) #unit*horizon and horizon*SampleSev sig

##interactions 3-way 
mod.u.h.ss.3.t <-lmer(percN~Unit*Horizon*SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.3.t) #no unit*horizon*samplesev

#from forward model selection appears that Unit*Horizon+Horizon*SampleSev+Trt is the best fit model

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) 
#model with lowest AIC: Unit*Horizon+Horizon*SampleSev+Trt (-136.2) 
#model with 2nd lowest AIC: Unit*Horizon*SampleSev+Trt (130.0) 
#3rd lowest AIC: Unit\*Horizon+SampleSev+Trt (-127.0)

#lowest BIC: Unit*Horizon+SampleSev+Trt (-79.3) 
#2nd lowest BIC: Unit*Horizon+Horizon\*SampleSev+Trt (-77.4) 
#3rd lowest BIC: Unit+Horizon (-74.7)

#%N model selected is Unit*Horizon+Horizon*SampleSev+Trt 
finalmod.N <- mod.u.h.2.ss.2.t 
rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)
```

### C stock model selection

```{r}
#Forwards model selection #Horizon 
mod.hor <- lmer(CStock ~ Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.hor) #

#Unit 
mod.unit <- lmer(CStock~Unit + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) #

#Soil subgroup 
mod.soil <- lmer(CStock~Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.soil) #

#Sample severity 
mod.sev <-lmer(CStock~SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.sev) #

#VegType 
mod.veg <- lmer(CStock~Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.veg) #

#Trt 
mod.trt <- lmer(CStock~Trt + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) # 

#unit and soil subgroup 
mod.u.sg <- lmer(CStock~Unit+Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.sg) #when accounting for unit, soil type 

#unit and veg type 
mod.u.v <- lmer(CStock~Unit+Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.v) #when accounting for unit, veg type 

#unit and horizon 
mod.u.h <-lmer(CStock~Unit+Horizon+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h) #

#unit, horizon, trt 
mod.u.h.t <- lmer(CStock~Unit+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t) #

#unit, horizon, samplesev 
mod.u.h.ss <- lmer(CStock~Unit+Horizon+SampleSev+ (1|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) #

#unit, horizon, samplesev, trt 
mod.u.h.ss.t <- lmer(CStock~Unit+Horizon+SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t) #

##interactions 2-way 
mod.u.h.2.ss.t <- lmer(CStock~Unit*Horizon+SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.t) #2-way interaction between unit and horizon 

mod.u.ss.2.h.t <- lmer(CStock~Unit*SampleSev+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.ss.2.h.t) #2-way interaction of unit and samplesev 

mod.u.t.2.ss.h <- lmer(CStock~Unit*Trt+SampleSev+Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.t.2.ss.h) #Unit*Trt

mod.u.h.ss.2.t <- lmer(CStock~Unit+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.2.t) #Horizon*SampleSev 

mod.u.h.t.2.ss <- lmer(CStock~Unit+Horizon*Trt+SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t.2.ss) #Horizon*Trt 

mod.u.h.ss.t.2 <- lmer(CStock~Unit+Horizon+SampleSev*Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t.2) #SampleSev*Trt

mod.u.h.2.ss.2.t <- lmer(CStock~Unit*Horizon+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.2.t) #unit*horizon and horizon*SampleSev 

##interactions 3-way 
mod.u.h.ss.3.t <-lmer(CStock~Unit*Horizon*SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.3.t) #unit*horizon*samplesev

#from forward model selection appears that ___________ is the best fit model

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) 
#model with lowest AIC: 
#model with 2nd lowest AIC: 
#3rd lowest AIC: 

#lowest BIC: 
#2nd lowest BIC:  
#3rd lowest BIC: 

#%N model selected is Unit*Horizon+Horizon*SampleSev+Trt 
finalmod.CStock <- mod.u.h.2.ss.2.t 
rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)

```

### N stock model selection

```{r}
#Forwards model selection #Horizon 
mod.hor <- lmer(NStock ~ Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.hor) #

#Unit 
mod.unit <- lmer(NStock~Unit + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) #

#Soil subgroup 
mod.soil <- lmer(NStock~Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.soil) #

#Sample severity 
mod.sev <-lmer(NStock~SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.sev) #

#VegType 
mod.veg <- lmer(NStock~Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.veg) #

#Trt 
mod.trt <- lmer(NStock~Trt + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) # 

#unit and soil subgroup 
mod.u.sg <- lmer(NStock~Unit+Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.sg) #when accounting for unit, soil type 

#unit and veg type 
mod.u.v <- lmer(NStock~Unit+Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.v) #when accounting for unit, veg type 

#unit and horizon 
mod.u.h <-lmer(NStock~Unit+Horizon+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h) #

#unit, horizon, trt 
mod.u.h.t <- lmer(NStock~Unit+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t) #

#unit, horizon, samplesev 
mod.u.h.ss <- lmer(NStock~Unit+Horizon+SampleSev+ (1|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) #

#unit, horizon, samplesev, trt 
mod.u.h.ss.t <- lmer(NStock~Unit+Horizon+SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t) #

##interactions 2-way 
mod.u.h.2.ss.t <- lmer(NStock~Unit*Horizon+SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.t) #2-way interaction between unit and horizon 

mod.u.ss.2.h.t <- lmer(NStock~Unit*SampleSev+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.ss.2.h.t) #2-way interaction of unit and samplesev 

mod.u.t.2.ss.h <- lmer(NStock~Unit*Trt+SampleSev+Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.t.2.ss.h) #Unit*Trt

mod.u.h.ss.2.t <- lmer(NStock~Unit+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.2.t) #Horizon*SampleSev 

mod.u.h.t.2.ss <- lmer(NStock~Unit+Horizon*Trt+SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t.2.ss) #Horizon*Trt 

mod.u.h.ss.t.2 <- lmer(NStock~Unit+Horizon+SampleSev*Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t.2) #SampleSev*Trt

mod.u.h.2.ss.2.t <- lmer(NStock~Unit*Horizon+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.2.t) #unit*horizon and horizon*SampleSev 

##interactions 3-way 
mod.u.h.ss.3.t <-lmer(NStock~Unit*Horizon*SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.3.t) #unit*horizon*samplesev

#from forward model selection appears that ___________ is the best fit model

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) 
#model with lowest AIC: 
#model with 2nd lowest AIC: 
#3rd lowest AIC: 

#lowest BIC: 
#2nd lowest BIC:  
#3rd lowest BIC: 

#%N model selected is Unit*Horizon+Horizon*SampleSev+Trt 
finalmod.NStock <- mod.u.h.2.ss.2.t 
rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)
```

### C:N ratio model selection

```{r}
#Forwards model selection #Horizon 
mod.hor <- lmer(CNratio ~ Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.hor) #

#Unit 
mod.unit <- lmer(CNratio~Unit + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) #

#Soil subgroup 
mod.soil <- lmer(CNratio~Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.soil) #

#Sample severity 
mod.sev <-lmer(CNratio~SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.sev) #

#VegType 
mod.veg <- lmer(CNratio~Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.veg) #

#Trt 
mod.trt <- lmer(CNratio~Trt + (1|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) # 

#unit and soil subgroup 
mod.u.sg <- lmer(CNratio~Unit+Soil_Subgroup + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.sg) #when accounting for unit, soil type 

#unit and veg type 
mod.u.v <- lmer(CNratio~Unit+Veg_Type + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.v) #when accounting for unit, veg type 

#unit and horizon 
mod.u.h <-lmer(CNratio~Unit+Horizon+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h) #

#unit, horizon, trt 
mod.u.h.t <- lmer(CNratio~Unit+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t) #

#unit, horizon, samplesev 
mod.u.h.ss <- lmer(CNratio~Unit+Horizon+SampleSev+ (1|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) #

#unit, horizon, samplesev, trt 
mod.u.h.ss.t <- lmer(CNratio~Unit+Horizon+SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t) #

##interactions 2-way 
mod.u.h.2.ss.t <- lmer(CNratio~Unit*Horizon+SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.t) #2-way interaction between unit and horizon 

mod.u.ss.2.h.t <- lmer(CNratio~Unit*SampleSev+Horizon+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.ss.2.h.t) #2-way interaction of unit and samplesev 

mod.u.t.2.ss.h <- lmer(CNratio~Unit*Trt+SampleSev+Horizon + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.t.2.ss.h) #Unit*Trt

mod.u.h.ss.2.t <- lmer(CNratio~Unit+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.2.t) #Horizon*SampleSev 

mod.u.h.t.2.ss <- lmer(CNratio~Unit+Horizon*Trt+SampleSev + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.t.2.ss) #Horizon*Trt 

mod.u.h.ss.t.2 <- lmer(CNratio~Unit+Horizon+SampleSev*Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.t.2) #SampleSev*Trt

mod.u.h.2.ss.2.t <- lmer(CNratio~Unit*Horizon+Horizon*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.2.ss.2.t) #unit*horizon and horizon*SampleSev 

##interactions 3-way 
mod.u.h.ss.3.t <-lmer(CNratio~Unit*Horizon*SampleSev+Trt+ (1|Plot_Name/Rep), data=df.noUB) 
anova(mod.u.h.ss.3.t) #unit*horizon*samplesev

#from forward model selection appears that ___________ is the best fit model

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) 
#model with lowest AIC: 
#model with 2nd lowest AIC: 
#3rd lowest AIC: 

#lowest BIC: 
#2nd lowest BIC:  
#3rd lowest BIC: 

#%N model selected is Unit*Horizon+Horizon*SampleSev+Trt 
finalmod.CNratio <- mod.u.h.2.ss.2.t 
rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)
```

##Final Models

```{r}
#%C final model 
finalmod.C <- mod.u.h.2.ss.2.t 
aov3C <- anova(finalmod.C) 
#unit, horizon, samplesev, unit*horizon & horizon*samplesev are sig predictors of %C

write.csv(aov3C, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Cmodel_aov3table.csv")

r.squaredGLMM(finalmod.C) 
#marginal (variance explained by fixed): 0.513; conditional (variance explained by both fixed and random): 0.562

#%N final model 
finalmod.N <- mod.u.h.2.ss.2.t 
aov3N <- anova(finalmod.N) 

r.squaredGLMM(finalmod.N) #marginal: 0.549; conditional: 0.615

write.csv(aov3N, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Nmodel_aov3table.csv")

#C stock final model
aov3Cstock <-anova(finalmod.CStock)

write.csv(aov3Cstock, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Cstockmodel_aov3table.csv")

r.squaredGLMM(finalmod.Cstock) #

#N stock final model
aov3Nstock <-anova(finalmod.NStock)

write.csv(aov3Nstock, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Nstockmodel_aov3table.csv")

r.squaredGLMM(finalmod.Nstock) #

#C:N final model
aov3CN <-anova(finalmod.CNratio)

write.csv(aov3CN, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/CNmodel_aov3table.csv")

r.squaredGLMM(finalmod.CNratio) #

```

###Contrasts on final models

```{r}
############ %C contrasts

###post hoc test on final model 
#fire history contrasts 
con.trt <- emmeans(finalmod.C, pairwise~Trt) 
t <- as.data.frame(con.trt$contrasts) %>% 
  mutate(., Unit=NA, Horizon=NA, SampleSev=NA) #save as a dataframe to merge with other contrasts

#burn severity contrasts 
con.samplesevbyhorizon <-emmeans(finalmod.C, pairwise~SampleSev|Horizon) 
con.samplesevbyhorizon #Mineral horizons no sig difference by burn severity #Organic horizons: 1&2 < 3&4 
ss.by.h <- as.data.frame(con.samplesevbyhorizon$contrasts) %>% 
  mutate(., SampleSev=NA, Unit=NA)

###what if allowed for 3-way interaction: 
#mod.u.h.ss.3.t <- lmer(percC~Unit*Horizon*SampleSev+Trt+(1|Plot_Name/Rep), data=df.noUB)
#test <- emmeans(mod.u.h.ss.3.t, pairwise~SampleSev|Unit+Horizon)
#test #mineral horizons no sig diff by burn severity; #O: BL no sig diff; H: 1 >2 & 1>4, but 1=3; WV: 2>3&4

con.horbyunitsamplesev <- emmeans(mod.u.h.2.ss.2.t, pairwise~Horizon|Unit+SampleSev) 
con.horbyunitsamplesev # O %C \> M %C except in burn sev class 4 in Butte Lake where O & M %C are equal
h.by.ss <- as.data.frame(con.horbyunitsamplesev$contrasts) %>%
  mutate(., Horizon=NA)

con.unitbyhorizon <-emmeans(mod.u.h.2.ss.2.t, pairwise~Unit|Horizon)
con.unitbyhorizon # Mineral: BL & H \<\< WV; Organic: BL \< H \< WV
u.by.h <- as.data.frame(con.unitbyhorizon$contrasts) %>% 
  mutate(., Unit=NA, SampleSev=NA)

#bind together relevant contrasts 
con.df <- rbind(ss.by.h, h.by.ss)
con.df <- rbind(con.df, u.by.h) 
con.df <- rbind(con.df, t)

write.csv(con.df, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Cmodel_contrasts.csv")
rm(con.horbyunitsamplesev, con.all.df, con.trt.df, con.unit.df, con.samplesevbyunithor, con.trt, con.unitbyhorizon, t, ss.by.u.h, h.by.ss, u.by.h)

C.con.df <- con.df 
rm(con.df)
```

```{r}
############ %N Contrasts

###post hoc test on final model 
con.trt <- emmeans(finalmod.N, pairwise~Trt) 
con.trt
t <- as.data.frame(con.trt$contrasts) %>%
  mutate(., Unit=NA, Horizon=NA, SampleSev=NA)

con.samplesevbyhor <-emmeans(finalmod.N, pairwise~SampleSev|Horizon)
con.samplesevbyhor #Mineral horizons no sig difference by burn severity
#Organic horizons: 2 \< 3&4; 1 is not different than 2, 3 or 4. This is strange. 
ss.by.h <- as.data.frame(con.samplesevbyhor$contrasts) %>%
  mutate(., SampleSev=NA, Unit=NA)

####what if allowed for 3-way interaction: 
#test <- emmeans(mod.u.h.ss.3.t, pairwise~SampleSev|Unit+Horizon) 
#test
#mineral horizons no sig diff by burn severity; #O: BL no sig diff; H: no sig diff; WV 2 \< 3 & 4 \

# what if included Unit*SampleSev interaction: 
#test <- lmer(percN~Unit*Horizon+Horizon*SampleSev+Unit*SampleSev+Trt + (1|Plot_Name/Rep), data=df.noUB) 
#test2 <- emmeans(test, pairwise~SampleSev|Unit+Horizon) 
#test2 #Error in crossprod(nbasis, x): "crossprod" is not a BUILTIN function

con.horbyunitsamplesev <- emmeans(finalmod.N, pairwise~Horizon|Unit+SampleSev) 
con.horbyunitsamplesev # O %N \> M %N # no difference in O & M %N: BL severity 1, BL severity 3, #H severity 3, BL severity 4, H severity 4
h.by.ss <- as.data.frame(con.horbyunitsamplesev$contrasts) %>% 
  mutate(., Horizon=NA)

con.unitbyhorizon <-emmeans(finalmod.N, pairwise~Unit|Horizon)
con.unitbyhorizon # Mineral: BL & H \<\< WV; Organic: BL \< H \< WV
u.by.h <- as.data.frame(con.unitbyhorizon$contrasts) %>% 
  mutate(., Unit=NA, SampleSev=NA)

#bind together relevant contrasts 
con.df <- rbind(ss.by.h, h.by.ss)
con.df <- rbind(con.df, u.by.h) 
con.df <- rbind(con.df, t)

write.csv(con.df, "C:/Users/iemus/Documents/UW_MS_LAVOSoils/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Nmodel_contrasts.csv")
rm(con.horbyunitsamplesev, con.all.df, con.trt.df, con.unit.df, con.samplesevbyunithor, con.trt, con.unitbyhorizon, t, ss.by.u.h, h.by.ss, u.by.h)

N.con.df <- con.df 
rm(con.df)
```

### Analyses including all samples (C above and N below detection)

```{r}
################ Z Analyses including all Samples (C above and N below detection)

################# Data Visualization

#visualize C/N data by treatment faceted by unit and horizon
ggplot(df.noUB, aes(Trt, percC)) +geom_boxplot() +
facet_grid(Horizon\~unit)

ggplot(df.noUB, aes(Trt, percN)) +geom_boxplot() +
facet_grid(Horizon\~unit)

#removing the samples with C and N concentrations outside of the
detection limits #of the EA analyzer ggplot(df.noUB, aes(Trt, cleanC))
+geom_boxplot() + facet_grid(Horizon\~unit)

ggplot(df.noUB, aes(Trt, cleanN)) +geom_boxplot() +
facet_grid(Horizon\~unit)

#C/N data by burn severity by unit and horizon ggplot(df.noUB,
aes(SampleSev, percC)) + geom_point() + facet_grid(Horizon\~unit)

ggplot(df.noUB, aes(SampleSev, percN)) + geom_point() +
facet_grid(Horizon\~unit)

################## Statistics

#linear mixed model with random effects of plot and core

########## %C Model Selection

#Forwards model selection #Horizon mod.hor \<- lmer(percC \~ Horizon +
(1\|Plot_Name/Rep), data=df.noUB) summary(mod.hor) anova(mod.hor)
#horizon is significant; p=2.2e-16

#Unit mod.unit \<- lmer(percC\~Unit + (1\|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) \# \*\*\* p=1.6e-8

#Soil subgroup mod.soil \<- lmer(percC\~Soil_Subgroup +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.soil) \# \*\*\* p=0.0002097

#Sample severity mod.sev \<-lmer(percC\~SampleSev + (1\|Plot_Name/Rep),
data=df.noUB) summary(mod.sev) anova(mod.sev) #\* p=0.030

#VegType mod.veg \<- lmer(percC\~Veg_Type + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.veg) #\*\*\* p=8.1e-6

#Trt mod.trt \<- lmer(percC\~Trt + (1\|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) \# Trt not sig p=0.5

#unit and soil subgroup mod.u.sg \<- lmer(percC\~Unit+Soil_Subgroup +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.sg) #when accounting for
unit, soil type is no longer a sig predictor

#unit and veg type mod.u.v \<- lmer(percC\~Unit+Veg_Type +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.v) #when accounting for
unit, veg type is no longer a sig predictor

#unit and horizon mod.u.h \<-lmer(percC\~Unit+Horizon+
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h) #both unit and horizon
sig predictors

#unit, horizon, trt mod.u.h.t \<- lmer(percC\~Unit+Horizon+Trt +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h.t) #trt not sig
predictor

#unit, horizon, samplesev mod.u.h.ss \<-
lmer(percC\~Unit+Horizon+SampleSev+ (1\|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) #burn severity is sig predictor

#unit, horizon, samplesev, trt mod.u.h.ss.t \<-
lmer(percC\~Unit+Horizon+SampleSev+Trt + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.ss.t) #trt still not sig predictor

##interactions 2-way mod.u.h.2.ss.t \<-
lmer(percC\~Unit*Horizon+SampleSev+Trt+ (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.2.ss.t) #2-way interaction between unit and
horizon (p=0.00048) mod.u.ss.2.h.t \<-
lmer(percC\~Unit*SampleSev+Horizon+Trt + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.ss.2.h.t) #no 2-way interaction of unit and
samplesev mod.u.t.2.ss.h \<- lmer(percC\~Unit*Trt+SampleSev+Horizon +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.t.2.ss.h) #no Unit*Trt
mod.u.h.ss.2.t \<- lmer(percC\~Unit+Horizon*SampleSev+Trt +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h.ss.2.t)
#Horizon*SampleSev mod.u.h.t.2.ss \<-
lmer(percC\~Unit+Horizon*Trt+SampleSev + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.t.2.ss) #no Horizon*Trt mod.u.h.ss.t.2 \<-
lmer(percC\~Unit+Horizon+SampleSev*Trt + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.ss.t.2) #no SampleSev*Trt

mod.u.h.2.ss.2.t \<- lmer(percC\~Unit*Horizon+Horizon*SampleSev+Trt +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h.2.ss.2.t) \#
unit*horizon and horizon*SampleSev sig

##interactions 3-way mod.u.h.ss.3.t
\<-lmer(percC\~Unit*Horizon*SampleSev+Trt+ (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.ss.3.t) #no unit*horizon*samplesev

#from forward model selection appears that
Unit*Horizon+Horizon*SampleSev+Trt is the best fit model
summary(mod.u.h.2.ss.2.t)

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg,
mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t,
mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss,
mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) #model with lowest
AIC: Unit*Horizon+Horizon*SampleSev+Trt (1831.3) #model with 2nd lowest
AIC: Unit*Horizon*SampleSev+Trt (1834.3) #3rd lowest AIC:
Unit\*Horizon+SampleSev+Trt (1834.7)

#%C model selected is Unit*Horizon+Horizon*SampleSev+Trt finalmod.C \<-
mod.u.h.2.ss.2.t rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg,
mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss,
mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h,
mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)

########## %N Model Selection

#Forwards model selection #Horizon mod.hor \<- lmer(percN \~ Horizon +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.hor) #horizon is
significant; p=2.2e-16

#Unit mod.unit \<- lmer(percN\~Unit + (1\|Plot_Name/Rep), data=df.noUB)
anova(mod.unit) \# \*\*\* p=2e-13

#Soil subgroup mod.soil \<- lmer(percN\~Soil_Subgroup +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.soil) \# \*\*\* p=0.1.38e-6

#Sample severity mod.sev \<-lmer(percN\~SampleSev + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.sev) #\* p=0.04812

#VegType mod.veg \<- lmer(percN\~Veg_Type + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.veg) #\*\*\* p=2.0e-8

#Trt mod.trt \<- lmer(percN\~Trt + (1\|Plot_Name/Rep), data=df.noUB)
anova(mod.trt) \# Trt not sig p=0.608

#unit and soil subgroup mod.u.sg \<- lmer(percN\~Unit+Soil_Subgroup +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.sg) #when accounting for
unit, soil type is no longer a sig predictor

#unit and veg type mod.u.v \<- lmer(percN\~Unit+Veg_Type +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.v) #when accounting for
unit, veg type is no longer a sig predictor

#unit and horizon mod.u.h \<-lmer(percN\~Unit+Horizon+
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h) #both unit and horizon
sig predictors

#unit, horizon, trt mod.u.h.t \<- lmer(percN\~Unit+Horizon+Trt +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h.t) #trt not sig
predictor

#unit, horizon, samplesev mod.u.h.ss \<-
lmer(percN\~Unit+Horizon+SampleSev+ (1\|Plot_Name/Rep), data=df.noUB)
anova(mod.u.h.ss) #burn severity is sig predictor

#unit, horizon, samplesev, trt mod.u.h.ss.t \<-
lmer(percN\~Unit+Horizon+SampleSev+Trt + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.ss.t) #trt still not sig predictor

##interactions 2-way mod.u.h.2.ss.t \<-
lmer(percN\~Unit*Horizon+SampleSev+Trt+ (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.2.ss.t) #2-way interaction between unit and
horizon (p=1.88e-6) mod.u.ss.2.h.t \<-
lmer(percN\~Unit*SampleSev+Horizon+Trt + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.ss.2.h.t) #no 2-way interaction of unit and
samplesev mod.u.t.2.ss.h \<- lmer(percN\~Unit*Trt+SampleSev+Horizon +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.t.2.ss.h) #no Unit*Trt
mod.u.h.ss.2.t \<- lmer(percN\~Unit+Horizon*SampleSev+Trt +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h.ss.2.t)
#Horizon*SampleSev mod.u.h.t.2.ss \<-
lmer(percN\~Unit+Horizon*Trt+SampleSev + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.t.2.ss) #no Horizon*Trt mod.u.h.ss.t.2 \<-
lmer(percN\~Unit+Horizon+SampleSev*Trt + (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.ss.t.2) #no SampleSev*Trt

mod.u.h.2.ss.2.t \<- lmer(percN\~Unit*Horizon+Horizon*SampleSev+Trt +
(1\|Plot_Name/Rep), data=df.noUB) anova(mod.u.h.2.ss.2.t) \#
unit*horizon and horizon*SampleSev sig

##interactions 3-way mod.u.h.ss.3.t
\<-lmer(percN\~Unit*Horizon*SampleSev+Trt+ (1\|Plot_Name/Rep),
data=df.noUB) anova(mod.u.h.ss.3.t) #no unit*horizon*samplesev

#from forward model selection appears that
Unit*Horizon+Horizon*SampleSev+Trt is the best fit model

anova(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg, mod.trt, mod.u.sg,
mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss, mod.u.h.ss.t, mod.u.h.2.ss.t,
mod.u.ss.2.h.t, mod.u.t.2.ss.h, mod.u.h.ss.2.t, mod.u.h.t.2.ss,
mod.u.h.ss.t.2, mod.u.h.2.ss.2.t, mod.u.h.ss.3.t) #model with lowest
AIC: Unit*Horizon+Horizon*SampleSev+Trt (-136.2) #model with 2nd lowest
AIC: Unit*Horizon*SampleSev+Trt (130.0) #3rd lowest AIC:
Unit\*Horizon+SampleSev+Trt (-127.0)

#lowest BIC: Unit*Horizon+SampleSev+Trt (-79.3) #2nd lowest BIC:
Unit*Horizon+Horizon\*SampleSev+Trt (-77.4) #3rd lowest BIC:
Unit+Horizon (-74.7)

#%N model selected is Unit*Horizon+Horizon*SampleSev+Trt finalmod.N \<-
mod.u.h.2.ss.2.t rm(mod.unit, mod.hor, mod.soil, mod.sev, mod.veg,
mod.trt, mod.u.sg, mod.u.v, mod.u.h, mod.u.h.t, mod.u.h.ss,
mod.u.h.ss.t, mod.u.h.2.ss.t, mod.u.ss.2.h.t, mod.u.t.2.ss.h,
mod.u.h.ss.2.t, mod.u.h.t.2.ss, mod.u.h.ss.t.2, mod.u.h.ss.3.t)

#############Final models \################ #%C final model finalmod.C
\<- mod.u.h.2.ss.2.t aov3C \<- anova(finalmod.C) #unit, horizon,
samplesev, unit*horizon & horizon*samplesev are sig predictors of %C

#write.csv(aov3C,
"C:/Users/iemus/Box/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Cmodel_aov3table.csv")
r.squaredGLMM(finalmod.C) #marginal (variance explained by fixed):
0.513; conditional (variance explained by both fixed and random): 0.562

#%N final model finalmod.N \<- mod.u.h.2.ss.2.t aov3N \<-
anova(finalmod.N) r.squaredGLMM(finalmod.N) #marginal: 0.549;
conditional: 0.615

#write.csv(aov3N,
"C:/Users/iemus/Box/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Nmodel_aov3table.csv")

########## Contrasts on final models

############ %C contrasts

###post hoc test on final model #fire history contrasts con.trt \<-
emmeans(finalmod.C, pairwise\~Trt) con.trt t \<-
as.data.frame(con.trt\$contrasts) %\>% mutate(., Unit=NA, Horizon=NA,
SampleSev=NA) #save as a dataframe to merge with other contrasts

#burn severity contrasts con.samplesevbyhorizon \<-emmeans(finalmod.C,
pairwise\~SampleSev\|Horizon) con.samplesevbyhorizon #Mineral horizons
no sig difference by burn severity #Organic horizons: 1&2 \< 3&4 ss.by.h
\<- as.data.frame(con.samplesevbyhorizon\$contrasts) %\>% mutate(.,
SampleSev=NA, Unit=NA)

###what if allowed for 3-way interaction: #mod.u.h.ss.3.t \<-
lmer(percC\~Unit*Horizon*SampleSev+Trt+(1\|Plot_Name/Rep), data=df.noUB)
#test \<- emmeans(mod.u.h.ss.3.t, pairwise\~SampleSev\|Unit+Horizon)
#test #mineral horizons no sig diff by burn severity; #O: BL no sig
diff; H: 1 \>2 & 1\>4, but 1=3; WV: 2\>3&4

con.horbyunitsamplesev \<- emmeans(mod.u.h.2.ss.2.t,
pairwise\~Horizon\|Unit+SampleSev) con.horbyunitsamplesev \# O %C \> M
%C except in burn sev class 4 in Butte Lake where O & M %C are equal
h.by.ss \<- as.data.frame(con.horbyunitsamplesev\$contrasts) %\>%
mutate(., Horizon=NA)

con.unitbyhorizon \<-emmeans(mod.u.h.2.ss.2.t, pairwise\~Unit\|Horizon)
con.unitbyhorizon \# Mineral: BL & H \<\< WV; Organic: BL \< H \< WV
u.by.h \<- as.data.frame(con.unitbyhorizon\$contrasts) %\>% mutate(.,
Unit=NA, SampleSev=NA)

#bind together relevant contrasts con.df \<- rbind(ss.by.h, h.by.ss)
con.df \<- rbind(con.df, u.by.h) con.df \<- rbind(con.df, t)

#write.csv(con.df,
"C:/Users/iemus/Box/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Cmodel_contrasts.csv")
rm(con.horbyunitsamplesev, con.all.df, con.trt.df, con.unit.df,
con.samplesevbyunithor, con.trt, con.unitbyhorizon, t, ss.by.u.h,
h.by.ss, u.by.h)

C.con.df \<- con.df rm(con.df)

############ %N Contrasts

###post hoc test on final model con.trt \<- emmeans(finalmod.N,
pairwise\~Trt) con.trt t \<- as.data.frame(con.trt\$contrasts) %\>%
mutate(., Unit=NA, Horizon=NA, SampleSev=NA)

con.samplesevbyhor \<-emmeans(finalmod.N, pairwise\~SampleSev\|Horizon)
con.samplesevbyhor #Mineral horizons no sig difference by burn severity
#Organic horizons: 2 \< 3&4; 1 is not different than 2, 3 or 4. This is
strange. ss.by.h \<- as.data.frame(con.samplesevbyhor\$contrasts) %\>%
mutate(., SampleSev=NA, Unit=NA)

####what if allowed for 3-way interaction: #test \<-
emmeans(mod.u.h.ss.3.t, pairwise\~SampleSev\|Unit+Horizon) #test
#mineral horizons no sig diff by burn severity; #O: BL no sig diff; H:
no sig diff; WV 2 \< 3 & 4 \# what if included Unit*SampleSev
interaction: #test \<-
lmer(percN\~Unit*Horizon+Horizon*SampleSev+Unit*SampleSev+Trt +
(1\|Plot_Name/Rep), data=df.noUB) #test2 \<- emmeans(test,
pairwise\~SampleSev\|Unit+Horizon) #test2 #Error in crossprod(nbasis, x)
: "crossprod" is not a BUILTIN function

con.horbyunitsamplesev \<- emmeans(finalmod.N,
pairwise\~Horizon\|Unit+SampleSev) con.horbyunitsamplesev \# O %N \> M
%N \# no difference in O & M %N: BL severity 1, BL severity 3, #H
severity 3, BL severity 4, H severity 4 h.by.ss \<-
as.data.frame(con.horbyunitsamplesev\$contrasts) %\>% mutate(.,
Horizon=NA)

con.unitbyhorizon \<-emmeans(finalmod.N, pairwise\~Unit\|Horizon)
con.unitbyhorizon \# Mineral: BL & H \<\< WV; Organic: BL \< H \< WV
u.by.h \<- as.data.frame(con.unitbyhorizon\$contrasts) %\>% mutate(.,
Unit=NA, SampleSev=NA)

#bind together relevant contrasts con.df \<- rbind(ss.by.h, h.by.ss)
con.df \<- rbind(con.df, u.by.h) con.df \<- rbind(con.df, t)

#write.csv(con.df,
"C:/Users/iemus/Box/MSWhitman/LAVO_FireHistory/Data/R/CarbonNitrogen/Nmodel_contrasts.csv")
rm(con.horbyunitsamplesev, con.all.df, con.trt.df, con.unit.df,
con.samplesevbyunithor, con.trt, con.unitbyhorizon, t, ss.by.u.h,
h.by.ss, u.by.h)

N.con.df \<- con.df rm(con.df)

####### Additional graphs from model building

#visualize 2-way interactions emmip(mod.u.h.2.ss.2.t, Unit \~ Horizon)
#estimated mean %C x axis Horizon colored by Unit
emmip(mod.u.h.2.ss.2.t, Horizon \~ SampleSev) #biggest change in
predicted %C occurred from low to moderate; particularly in organic
horizon
```
